Inefficient Concessions and Mediation

### 2017-09-12 draft.tex
turn theorem:3 into lemma:efficiency

\begin{center}
\captionof{table}{Stage Game Payoffs}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& T, T& -D, T+W \\ \hline
	Fight & T+W,-D& W-D, W-D \\ \hline
\end{tabular}
\end{center}

@@

\section{Material Value of Concession and Inefficiency} 
\label{sec:IC}
We now model concessions to potentially have real material help or harm to the giver. The interpretation is that the country which receives the concessions can use them to build up their military or to build up their civil society. The former hurts the giver of the concessions in war; the latter helps the giver in peace. The Golan heights is such an example: if Syria is an ally of Israel, the Golan Heights is a reasonable concession to make, but if Syria later turns out to betray a trust, surrender of the strategic land would be disastrous for Israel. %Another example is ADD more examples

@@

\begin{center}
\captionof{table}{Stage Game Payoffs when Concessions have Material Value}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2 g_1)$, & \small$-D(m_2+(1-\alpha_2) g_1),$ \\
  & \small$T(s_1+\alpha_1 g_2)$ & \small$T(s_1+\alpha_1 g_2)  +W(m_2+(1-\alpha_2) g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2 g_1)+W(m_1+(1-\alpha_1) g_2),$& \small$W(m_1+(1-\alpha_1) g_2)-D(m_2+(1-\alpha_2) g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1) g_2) $ & \small$W(m_2+(1-\alpha_2) g_1)-D(m_1+(1-\alpha_1) g_2)$ \\ \hline
\end{tabular}
\end{center}
\vskip.1in
\begin{theorem}
	Assume the high types are not patient enough to separate without concessions. There are parameters under which the ability of the low-type country to use concessions to reduce the payoffs of its negotiating partner during a 'fight' stage (i.e. $\al_L < 1$) destroys the negotiating partners' ability to separate through concessions, i.e. to achieve peace.
	\label{theorem:matval}
\end{theorem}

@@

@@

We now relax the assumption about the benefit of a concession being equal to its cost and instead allow countries to ``burn'' a portion of the concession they give. For example, an expensive display of pageantry to a dignitary might benefit the receiving country very little, if at all, but it still signals good intentions on behalf of the giving country. The option to give efficient concessions is still available along with new options to give a variety of less efficient concessions.

In the model, countries choose a scalar $e$ to multiply by their given concession $g$. $0 \leq e \leq 1$. As before, $C(g)= g$ but while the receiving country observes the full concession $g$, its benefit is now only $eg$.\footnote{In a more general model not relevant to results presented here, we could also think of $e$ as a function $E(\cdot)$. The argument is still $g$, but instead of a scalar, $E(\cdot)$ can transform $g$ into any concession, not only a numerical value. These gifts will be converted into utilities by cost and benefit functions, which then allow the possibility of a gift having efficiency greater than 1. The transformation function $E(\cdot)$ can be thought of as the type of gift, and has a cost to the sender and a direct benefit to the receiver in addition to its signaling value.} From here on in the analysis, we use $g_l=0$ and $g_h=g_h^*$ to denote the equilibrium concenssions as presented in the previous section and calculated in the appendix. \\
\\

\begin{center}
\captionof{table}{Stage Game Payoffs when Concessions have Material Value}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2 e g_1)$, & \small$-D(m_2+(1-\alpha_2)e g_1),$ \\
  & \small$T(s_1+\alpha_1 e g_2)$ & \small$T(s_1+\alpha_1 e g_2)  +W(m_2+(1-\alpha_2)e g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2 e g_1)+W(m_1+(1-\alpha_1) e g_2),$& \small$W(m_1+(1-\alpha_1)e g_2)-D(m_2+(1-\alpha_2)e g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1) e g_2) $ & \small$W(m_2+(1-\alpha_2)e g_1)-D(m_1+(1-\alpha_1)e g_2)$ \\ \hline
\end{tabular}
\end{center}

\begin{theorem}
	It is optimal to give efficient gifts (where $e=1$) in a separating equilibrium.
	\label{theorem:3}
\end{theorem}
\emph{Proof:}
The benefit of a gift appears on both sides of the IC for both types and thus cancels out in the low type IC $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type IC:  $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$.\footnote{Elaboration and explanation of the IC constraints is in the proof of Theorem~\ref{theorem:2}.} Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types. However, the expected separation utilities under this game are now
\be
U_h= pX_{TT}+(1-p)X_{FF} - g_h+peg_h 
\ee
and
\be
U_l=X_{FF}+peg_h 
\ee
 $\frac{dU}{de}>0$ for both countries so it is never optimal have $e<1$.\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} \hfill $\blacksquare$

>>>>

\begin{center}
\captionof{table}{Stage game payoffs}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& T, T& -D, T+W \\ \hline
	Fight & T+W,-D& W-D, W-D \\ \hline
\end{tabular}
\end{center}

@@

\section{Material Value of Concession and Inefficiency} 
\label{sec:IC}

\subsection{Concessions with Material Value}
We now model concessions to potentially have real material help or harm to the giver. The interpretation is that the country which receives the concessions can use them to build up their military or to build up their civil society. The former hurts the giver of the concessions in war; the latter helps the giver in peace. The Golan heights is such an example: if Syria is an ally of Israel, the Golan Heights is a reasonable concession to make, but if Syria later turns out to betray a trust, surrender of the strategic land would be disastrous for Israel. %Another example is ADD more examples

@@

\begin{center}
\captionof{table}{Stage game payoffs when concessions have material value}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2 g_1)$, & \small$-D(m_2+(1-\alpha_2) g_1),$ \\
  & \small$T(s_1+\alpha_1 g_2)$ & \small$T(s_1+\alpha_1 g_2)  +W(m_2+(1-\alpha_2) g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2 g_1)+W(m_1+(1-\alpha_1) g_2),$& \small$W(m_1+(1-\alpha_1) g_2)-D(m_2+(1-\alpha_2) g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1) g_2) $ & \small$W(m_2+(1-\alpha_2) g_1)-D(m_1+(1-\alpha_1) g_2)$ \\ \hline
\end{tabular}
\end{center}
\vskip.1in
\begin{theorem}
	Assume the high types are not patient enough to separate without concessions. There are parameters under which the low-type country's ability to use concessions to reduce the payoffs of its negotiating partner during a `fight' stage (i.e. $\al_L < 1$) destroys the negotiating partners' ability to separate through concessions, i.e. to achieve peace when there are two high types.
	\label{theorem:matval}
\end{theorem}

@@

\subsection{Material Value of Concession can be Destroyed}
We now relax the assumption about the benefit of a concession being equal to its cost and instead allow countries to ``burn'' a portion of the concession they give. For example, an expensive display of pageantry to a dignitary might benefit the receiving country very little, if at all, but it still signals good intentions on behalf of the giving country. The option to give efficient concessions is still available along with new options to give a variety of less efficient concessions.

In the model, countries choose a scalar $e$ to multiply by their given concession $g$. $0 \leq e \leq 1$. As before, $C(g)= g$ but while the receiving country observes the full concession $g$, its benefit is now only $eg$.\footnote{In a more general model not relevant to results presented here, we could also think of $e$ as a function $E(\cdot)$. The argument is still $g$, but instead of a scalar, $E(\cdot)$ can transform $g$ into any concession, not only a numerical value. These gifts will be converted into utilities by cost and benefit functions, which then allow the possibility of a gift having efficiency greater than 1. The transformation function $E(\cdot)$ can be thought of as the type of gift, and has a cost to the sender and a direct benefit to the receiver in addition to its signaling value.} From here on in the analysis, we use $g_l=0$ and $g_h=g_h^*$ to denote the equilibrium concenssions as presented in the previous section and calculated in the appendix.

\begin{center}
\captionof{table}{Stage game payoffs when concessions have material value that can be destroyed}
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2 e g_1)$, & \small$-D(m_2+(1-\alpha_2)e g_1),$ \\
  & \small$T(s_1+\alpha_1 e g_2)$ & \small$T(s_1+\alpha_1 e g_2)  +W(m_2+(1-\alpha_2)e g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2 e g_1)+W(m_1+(1-\alpha_1) e g_2),$& \small$W(m_1+(1-\alpha_1)e g_2)-D(m_2+(1-\alpha_2)e g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1) e g_2) $ & \small$W(m_2+(1-\alpha_2)e g_1)-D(m_1+(1-\alpha_1)e g_2)$ \\ \hline
\end{tabular}
\end{center}

Before turning to the question of how the possibility of destroying the value of concessions impacts the ability of the parties to separate, we first formally establish the welfare implications.

\begin{lemma}
	Assuming that a separating equilibrium exists when concessions have material value and are efficient (where $e=1$), it is optimal to give efficient gifts in a separating equilibrium.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:}
The benefit of a gift appears on both sides of the IC for both types and thus cancels out in the low type IC $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type IC:  $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$.\footnote{Elaboration and explanation of the IC constraints is in the proof of Theorem~\ref{theorem:2}.} Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types. However, the expected separation utilities under this game are now
\be
U_h= pX_{TT}+(1-p)X_{FF} - g_h+peg_h 
\ee
and
\be
U_l=X_{FF}+peg_h 
\ee
 $\frac{dU}{de}>0$ for both countries so it is never optimal have $e<1$.\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} \hfill $\blacksquare$

@@
